### 2024-05-06 main.tex
Revision

%                                                                 aa.dem
% AA vers. 9.1, LaTeX class for Astronomy & Astrophysics
% demonstration file
%                                                       (c) EDP Sciences
%-----------------------------------------------------------------------
%
%\documentclass[referee]{aa} % for a referee version
%\documentclass[onecolumn]{aa} % for a paper on 1 column  
%\documentclass[longauth]{aa} % for the long lists of affiliations 
%\documentclass[letter]{aa} % for the letters 
%\documentclass[bibyear]{aa} % if the references are not structured 
%                              according to the author-year natbib style

%
\documentclass{aa}

%
\usepackage{graphicx}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{txfonts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\usepackage[options]{hyperref}
% To add links in your PDF file, use the package "hyperref"
% with options according to your LaTeX or PDFLaTeX drivers.
%
\begin{document}

   \title{The variability  of Betelgeuse explained by surface convection}

    \author{{ Q.~Pilate}\inst{1},{ A.~L{\'o}pez Ariste}\inst{1},{ A.~Lavail}\inst{1},{ Ph. Mathias}\inst{1} }

   \institute{IRAP, Universit\'e de Toulouse, CNRS, CNES, UPS.  14, Av. E. Belin. 31400 Toulouse, France 
             }

   \date{Received ...; accepted ...}

% \abstract{}{}{}{}{} 
% 5 {} token are mandatory

  \abstract

   \keywords{
               }

   \maketitle
%
%-------------------------------------------------------------------

\section{Introduction}
Betelgeuse is a prototypical red supergiant (RSG) known to be a semi-regular variable. Several periods for this variability can be found in the 
literature, usually clustered around the so-called long secondary period (LSP) of 2000 days, plus shorter periods of 400 and 200 days \citep{kiss_variability_2006}. 
The origin of the LSP remains poorly known: some have tried to find the explanation of  the LSP in the lifetime of the large convective cells present at the surface of RSG (e.g \cite{stothers_giant_2010}) while others have invoked a the magnetic field \citep{wood_long_2004}. 
The shorter periods of 400 and 200 days seem to have a different origin than the LSP. By studying a large sample of 
RSG, \cite{kiss_variability_2006} attributed the 400 days period to radial pulsation modes, while the 200 days period was attributed to the first overtone \citep{joyce_standing_2020}, 
 in agreement with the historical work of \cite{stothers_pulsation_1969}.\\

Besides radial pulsation, surface activity leading to random brightness variation has also been mentioned to explain the variability of Betelgeuse. 
\textbf{Having been firstly theorised by \cite{schwarzschild_scale_1975} and \cite{gray_mass_2008} 
proposed that the 400 days period was a consequence of the convective activity}. Betelgeuse is known to present large convective cells with lifetimes in the order of one to two years \citep{lopez_ariste_convective_2018} and measurable changes 
in the span of one week. Bright convection cells near disk center will increase the integrated brightness of the star when 
compared with other situations where such cells are found near the edges. So even without calling for the formation of 
dust or other changes in opacity, one may argue that the simple evolution of convective patterns of the star may create a variability. Such 
variability could be expected to be random, but will present quasi-periodicities related to the typical time scales of the 
convective patterns \citep{gray_mass_2008}. \\

%more or less well defined 
%periods around 400 and 200 days. 
%Such periods are often linked to radial pulsation modes,   and have been used to model 
%the characteristics of the resonant cavity, hence the size of the star and eventually its evolutionary stage \citep{saio_evolutionary_2023}.

At the end of 2019, Betelgeuse reached a historical minimum in its luminosity, called the Great Dimming \citep{guinan_fall_2020}. 
From interferometric and spectroscopic data, it has been determined that this event was caused by a cloud of dust formed close to the line of 
sight \citep{montarges_dusty_2021}. 
Interferometric images show a drop in luminosity in the southern hemisphere of Betelgeuse, that might be caused by a 
mass loss event and leading to this dimming \citep{dupree_great_2022}. Other hypothesis have  also been explored , such as a drop in temperature \citep{harper_photospheric_2020}, or an increase in the molecular opacity \citep{kravchenko_atmosphere_2021}. 
This event is an extreme illustration of  changes in brightness that were not due to any kind of pulsation but to a change on the 
brightness distribution over the stellar disk.
%due to the presence of large quantities of dust.
Since the end of the Great Dimming event, Betelgeuse has continued its random variation of brightness. 
But quite interestingly, 
it has been shown by \cite{jadlovsky_analysis_2023} and \cite{dupree_great_2022} that the periodicity of Betelgeuse has changed since the dimming. Using the light curves from AAVSO, 
the cited authors showed that before the dimming, the dominant period of Betelgeuse was the 400 days period, while after the great dimming the main periods of Betelgeuse have shortened, oscillating between 97 days and 230 days \citep{dupree_great_2022}, revealing a change in the behaviour of the atmosphere of Betelgeuse.
While the great dimming may be seen as 
a singular event, such modifications of the variability periods  bring up the question whether all changes in brightness cannot be due to similar changes 
in the brightness  distribution over the disk, and unrelated to any pulsation phenomenon. 
%Our interest in an alternative explanation of the variability
%in terms  of convective patterns arose. 
In order to address this question we  seek for the typical periods of 400 and 200 days in  
observational proxies which are related to the convective activity but not to any pulsation, such as the linear polarization spectra.\\
%but unrelated to pulsations and which cannot be directly linked to brightness.
%Faced with the dramatic change in stellar parameters and age that such change in a pulsation period would 
%require,

Linear polarization in the atomic lines of the spectrum of Betelgeuse, discovered by \cite{auriere_discovery_2016}, has been interpreted as the joint action of 
two mechanisms. First the depolarization of the continuum by atoms which absorb linearly polarized light from the continuum and re-emit unpolarized light, the 
continuum photons being polarized by Rayleigh scattering. This depolarization produces signals with an azimuth symmetry over the disk that 
would  cancel out the net linear polarization on a homogeneous disk. Thus, such mechanism of production of polarization must be combined with an inhomogenous disk to produce the net linear polarization 
signal observed. This interpretation suggested the possibility of mapping those brightness inhomogeneities. This has been done by \cite{lopez_ariste_convective_2018} 
even producing 3-dimensional images of the atmosphere of Betelgeuse \citep{lopez_ariste_three-dimensional_2022} when taking advantage of the different heights of formation of 
different lines in the spectrum of Betelgeuse. The images produced compare very well with contemporaneous images made with interferometric 
techniques \citep{montarges_close_2016} and show clear convective patterns, akin to solar granulation.

Linear polarization observed in the atomic lines is therefore a proxy of convection, unrelated to radial 
pulsations but linked to the brightness inhomogeneities due to the convective patterns in Betelgeuse. If one can find the 
aforementioned variability periods 
in linear polarization, we must conclude that these periods are related to the convective activity which originates the linear 
polarization signals.
%and unrelated to any radial pulsation phenomena. 
This is the purpose of the present work.

%In section \ref{Section 2} we describe the dataset of linear polarization obtained with the Telescope Bernard Lyot (TBL) and the 
%Lomb-Scargle technique used to derive periods. 
%In section \ref{section 3} we associate the 200 days period to convective timescale of the smallest granules and the 330 days periodicity to the largest granules, and we provide an explanation in the change of variability of Betelgeuse before and after the great dimming. 
%In section \ref{Conclusion}, we summup our work.

\section{Searching for periodicities in the polarization spectra of Betelgeuse}
\label{Section 2}

Betelgeuse has been observed for the last 10 years with Narval and Neo-Narval at the Telescope Bernard Lyot. These two instruments measure the polarization 
over the visible spectra of Betelgeuese (390-1000 nm) with high spectral resolution (R=65000) and high polarimetric sensitivity. Despite 
this high sensitivity, the signal-to-noise ratios per spectral bin are not sufficient to measure the weak polarization signals in individual 
atomic lines. These amplitudes are today known to be of the order of $10^{-4}$ times the continuum intensity, and only exceptionally do they 
reach amplitudes of $10^{-3}$ the continuum. In these exceptional observations, enough photons can be accumulated per spectral bin to 
see the linear polarization signal above noise \citep{auriere_discovery_2016}. But most commonly, the amplitudes of linear polarization are 
below noise levels and measuring them requires the addition of the signals of thousands of lines to reduce  noise and increase the 
signal-to-noise ratios. This line addition 
is done through a technique called Least-Squares Deconvolution (LSD) \citep{donati_spectropolarimetric_1997} and assumes that the 
linear polarization signal is similar 
in all spectral lines up to scale factors in the sampling and amplitude. This technique has been successfully used in the past to measure 
magnetic field distributions over stellar surfaces and it is now used to produce images of the brightness distributions in the photosphere 
of Betelgeuse and other red supergiants as cited in the Introduction.

In this work we shall look into the signals produced through these line-addition techniques, and which produce a pseudo-line in intensity 
and linear polarization which does not belong to any atomic species in particular but to an average of all of them present and emitting in the 
photosphere of Betelgeuse. After such procedures of line addition, these signals carry the coherent signals present in the photospheric lines, but the particularities of this or 
that spectral line are erased. All these data has been presented before by \cite{auriere_discovery_2016}, \cite{mathias_evolution_2018} and 
\cite{lopez_ariste_three-dimensional_2022}, which also give details on observing times and conditions as well as more detailed description 
on the data reduction\footnote[1]{Beyond a 2-year proprietary embargo, and up to technical issues, all these data is available through 
PolarBase (http://polarbase.irap.omp.eu/).}.

%\subsection{Lomb-Scargle periodograms of the profiles}

\subsection{The variability in the Stokes profiles}

Trying to find periods in an astrophysical context carries the hurdle of unevenly spaced observed data. To overcome this issue, 
we used  Lomb-Scargle periodograms \citep{lomb_least-squares_1976,scargle_studies_1982}. For a given set of data, one fits sinusoids to the data
using least-squares, with frequencies sampled between the first and the last observation. The better the fit, the higher the power attributed to a frequency. 
We applied this technique to the  LSD profiles of Stokes $I$, $U$, $Q$ and to the total linear polarization of Betelgeuse observed by the TBL during 
that span of 10 years.

We first computed the Lomb-Scargle periodogram at each wavelength of the LSD profile, between -25 km/s and +60 km/s to span the entire 
signal present in the LSD profile. \cite{lopez_ariste_convective_2018} found the most blueshifted signal in linear polarization  towards -20 km/s, 
and interpreted it as  the maximum velocity of the rising plasma, seen from  Earth. The most redshifted signal was often found at +40 km/s. 
This velocity is interpreted
as the rest velocity of the star in our reference frame. Seldom, signals at velocities larger than +40 km/s can be found 
and are interpreted as either dark and cold plasma falling back to Betelgeuse, 
or to rising plumes of plasma located in the hidden face of Betelgeuse,
and rising so high that they appear above the limb.

% The LS periodogram  computed for each frequency 
% in the LSD profile must show periodicities around the LSP, but also on the 400 days and 200 days period. If the LS periodogram finds a periodicity at a
% given wavelength, it would appear as a strong LS power. However, we expect a change in the periodicity after the dimming. 
% This would lead to a decrease in the Lomb-Scargle power. By comparing LS periodogram before and after the dimming, we expect to see a change in the LS power,
% especially at the 400 days period.

Figures \ref{LS intensity},\ref{LS Q},\ref{LS U} and \ref{LS linear polarization} show the Lomb-Scargle periodogram of the LSD profiles of respectively 
the intensity, Stokes $Q$, Stokes $U$ and the total linear polarization. 
%The abscissa is the frequency (in $\mathrm{days^{-1}}$), while the ordinate is the heliocentric radial velocity in km/s. 
For each figure, the uper panel is the Lomb-Scargle periodogram of each velocity bin (black line) and the total average (red line). The green line corresponds to the window function. 
The lower panel is the Lomb-Scargle periodogram represented for each velocity bin, while the white dotted lines mark the 400 and 200 days periods for 
reference. The right panel shows the mean profile of each Stokes parameter at each velocity bin.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Intensity.pdf}
    \caption{Lomb-Scargle periodogram of the LSD profile of intensity.
    The upper panel is the Lomb-Scargle periodogram for each velocity bin (black line) and the average (red line). The green line is the window function.
    In the lower pannel, the two white dashed lines mark respectively the 400 days and the 200 days periods. The right panel is the average intensity profile for each velocity bin.}
    \label{LS intensity}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes Q.pdf}
    \caption{Same as figure \ref{LS intensity} for Stokes $Q$. }
    \label{LS Q}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Stokes U.pdf}
    \caption{Same as figure \ref{LS intensity} for Stokes $U$.}
    \label{LS U}
\end{figure}

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle linear polarization.pdf}
    \caption{Same as figure \ref{LS intensity} for total linear polarization: $\sqrt{Q^2+U^2}$.}
    \label{LS linear polarization}
\end{figure}

%\section{Recovering the variability of Betelgeuse in the LSD profile}

%\subsection{Intensity}
Looking into Fig. \ref{LS intensity}, which shows the Lomb-Scargle periodogram of the LSD of the intensity profile, 
we notice that the 400 and 200 days periods seem to be recovered by the periodogram. However, the 400 days period is uncomfortably close to 
the peak of the window function. We also notice that the two periods are present at the same HRV. Interestingly, 
\cite{mathias_evolution_2018} already found this 200 days periodicity in spectroscopic observations, in spite of a shorter observation period.

Figures \ref{LS Q} and \ref{LS U} show the Lomb-Scargle periodograms of Stokes $Q$ and $U$. These periodograms are completely different 
from the intensity one. A strong signal is present in fig. \ref{LS Q} around 200 days. Concerning Stokes U, Fig. \ref{LS U}, it appears on the other hand that the main period is around $0.003 \ \mathrm{days^{-1}}$ (330 days). 
Other periods such as those at 200 days or 250 days ($0.004 \ \mathrm{days^{-1}}$) are present but difficult to trust. From both Stokes $Q$ and $U$, we recover the 200 days period, and also a 330 days period which  is more or less close to the 400 days period found in the literature. 
Figure \ref{LS linear polarization} shows the Lomb-Scargle periodogram of $\sqrt{Q^2+U^2}$, the total linear polarization of Betelgeuse. 
The periodogram recovers the strong powers at both 330 days and 200 days, already found in either Stokes $Q$ and $U$.

%Nonetheless, the power of this period peaks at a HRV of around 40 km/s. We recall that this velocity is the rest velocity of the star. 
%Every signal beyond this velocity is either due to  plasma sinking in Betelgeuse, or to plumes of plasma rising beyond and above the limb of the star. 
%Such plumes from the back hemisphere have been often seen in other red supergiants like $\mu$ Cep \citep{lopez_ariste_height_2023}. 
%This suggests that both signals at these period of 400 days found in intensity and total linear polarisation 
%could well be due to the periodic appearance of plumes in the back hemisphere rising high enough to appear above
%the visible limb and generating a supplementary source of brightness. Once again we find a plausible explanation 
%of these periods unrelated to pulsations, but rather to the convective dynamics of the star.

% Up to this point, we want to pinpoint the two elements that are a clue 
% to understand the variability of Betelgeuse: the 400 days period is present in the linear polarization signal and also at a HRV above 40 km/s.
% In any case, from the linear polarization and the intensity profile, we recover the different periods of Betelgeuse. Since linear polarization is 
% associated to convection and because we recover the periods of Betelgeuse, the main mechanism behind the variability of Betelgeuse points toward surface convection.

\subsection{The variability in polarimetric imaging}

Using linear polarization, \cite{lopez_ariste_convective_2018} were able to reconstruct images of Betelgeuse that have been successfully compared 
to inteferometric images of \cite{montarges_close_2016}. The images are produced by finding the brightness distribution that better fits the observed linear 
polarization LSD profiles, using a Marquardt-Levenberg minimisation.

Betelgeuse is observed in average every month by the TBL, and hence
we are able to follow its surface activity through this technique of polarimetric imaging. This has  allowed in the past the estimation of  
the size and the lifetime of the convective cells
at the surface \citep{lopez_ariste_convective_2018}. From these images, we can also compute a photo-center, which is sensitive to the size and the number of convective cells. 
An homogeneous star will have a photo-center displacement coinciding with the barycenter of the star, whereas a star with one or two large convective cells will have a
photo-center displacement more important, up to a few percent of the stellar radius in the case of \textbf{red giants?} \citep{chiavassa_probing_2022}.

Since the photo-center is linked to surface convection, it is worth checking for periods in its 
dynamics over the 10 years of observations of Betelgeuse at the TBL. Such periods will not 
be completely unrelated to the ones presented in the previous section, but will capture 
other aspects of the phenomena at work.

%We computed the Lomb-Scargle periodogram of the displacement of the photo-center.
But, before proceeding to search for periods in the movements of the photo-center, we have to warn about the interpretation of such images. 
Linear polarization suffers from a $180 ^\circ$ degree ambiguity already mentioned in \cite{auriere_discovery_2016}. For that reason, our 
images can be rotated by $180^\circ$ degrees, and the brightness distribution will still fit the observed LSD polarization profiles. 
If a particular observation were to be treated independently of previous observations, the solution found by 
the algorithm could be any of the ambiguous solutions possible. To obtain a continuity 
between the time series of images, for a given day we use the brightness distribution of the previous day as initial point 
of the  fitting iteration. The first image of the series  starts its minimization iteration with 
random brightness distribution. 
Even though we are able to produce a series of images  consistent with each other, it is important 
to keep in mind that the photo-center displacement computed from the series 
will be affected by that first image of the series. 
To overcome this issue, we decided to compute
the photo-center displacement from 100 different time serieas, each one starting from a different initial image, in an effort to recover 
from such ensemble
properties that are independent of the choice of the first image.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Lomb-Scargle Photo-center.pdf}
    \caption{Lomb-Scargle periodogram of the 100 series of photo-center displacements of Betelgeuse. Each blue line correspond to a Lomb-Scargle periodogram of one series of photo-center
     displacement. The black line is the average of the 100 periodograms. The red dashed lines mark the 400 and 200 days periods respectively 
     (from left to right). As before, the green line is the window function.}
    \label{LS photocenter}
\end{figure}

Figure \ref{LS photocenter} shows the Lomb-Scargle periodogram of the photo-center displacement of each one of those 100 series (blue lines) and the average Lomb-Scargle periodogram (black line), 
%As before, the upper panel shows the Lomb-Scargle periodogram from data before the great dimming only, whereas we used the whole set of data for the lower panel.
while the red dashed lines mark the 400 and 
200 days periods respectively. As before, the window function is represented by the green line. We find two bumps around the 400 days period, which are not 
sufficiently significant to confirm the presence of such period on their own. However, they concur with the results of the previous section, 
and since polarimetry imaging involves only surface convection, this is a further argument towards a variability explained by surface convection alone. 
%First, we find that the Lomb-Scargle power is high around the LSP, though
%uncomfortably close to the auto-correlation peak.
%We also notice a decrease in the Lomb-Scargle power from  before to after the dimming, in 
%concurrence to what was seen in the periodograms of the LSD profiles. Before it, there are two smalls bumps around the 400 days period. 
%But those bumps almost completely disappear after the dimming. Even though it is 
%hard to identify the 400 days period from those bumps alone,
%we still find that there is a difference before and after the dimming. 
%And this is concluded  from images that exclusively involve surface convection.
%The Lomb-Scargle periodogram of the photo-center displacement points towards a variability explained by convection only,
%and this is consistent with what we obtained from linear polarization.
%Both the photo-center displacement and the LS of the linear polarization point toward a variability explained by surface convection.

\section{Explanations for the variability of the light curve of Betelgeuse}
\label{section 3}
%Before summing up the conclusions of this work, we can afford proposing alternative scenarios for the observed periodicities in the light curve, 
%the LSD profiles and the displacement of the photo-center of the inferred photospheric image reconstructions. 
The presence of the same 
periods and, after taking into consideration the time span and sparsity of the TBL data series, the larger amplitude of the peaks found in
polarization data  compared to those  in the light curve point to a common origin, one that is rather based upon convective dynamics.

\textbf{The 200-days period appears prominently in the Stokes Q and U profiles. We can }
associate the 200-days period to the typical timescales on which the Stokes parameters evolve. 
The LSD profiles of Stokes $Q$ or $U$ will slightly change when we observe Betelgeuse every month, but are completely overhauled 
on time scales of several months. The Lomb-Scargle periodograms appear to capture these dynamics at periods of about 200 days.
 Such changes in the profiles are directly linked to the dynamics of the surface,  to  the evolution and movements 
of convective cells over the visible hemisphere. So the 200-day period appears to be related to the evolution of the convective patterns 
on the surface of Betelgeuse.  The fact that this peak is strongly present in Stokes $Q$ but less in $U$ 
can be interpreted as a temporal situation of how the convective patterns have been distributed in the last 10 years.
%points toward a more stochastic origin, which can fully be explained by convection.\

The other period found in the total linear polarization and Stokes $U$, the 330-days period, is more or less close to the 400 days period 
often present in the literature (e.g \cite{kiss_variability_2006} found a period $388 \pm 30$ days). It has been found
by \cite{lopez_ariste_convective_2018} that large convective cells can live  one  to two years, and this period could be associated to the 
timescales on which the largest granules evolve. This typical timescale has also been observed in the numerical simulations and might 
play an important role in the determination of stellar distances through the displacement of the photo-center \citep{chiavassa_probing_2022}. 
Once again we observe this 330-days periodicity to be more present in Stokes $U$ than in $Q$, what may unravel a 
random situation due to convective motions in the last 10 years. This interpretation appears to be  consistent with the work of \cite{gray_mass_2008}, 
who attributed the 400 days period to the large convective cells.

\textbf{Finally, the intensity profiles do  seem to vary in  400 days periods, but the measured peak is uncomfortably close to a 
peak of the window function of the time series observed with the TBL. This 400-days period, it could be a mix of the typical convective 
timescales of large granules associated with stochastic pulsations exited by convective cavities. Up to this point, the model developed to 
interpret linear polarization does not involve pulsations, that is why we interpret the 200 days period as the convective turnover timescale. 
The 400 days period involve the intensity profile, thus the 400 days period is probably caused by both convective timescales and stochastic 
pulsations.}
 %As showed by
%\cite{lopez_ariste_convective_2018}, some of those convective cells can live up for years. We may speculate
%that the LSP may be related to those long living structures. However the present span of data from the TBL is insufficient to do more than
%just speculate. It should be noticed that  
% the peak is not present in stokes U, and this may point to a more stochastic event than just long-living structures.

\subsection{The light curve of Betelgeuse from the point of view of convection}

%After examining the periods found by the Lomb-Scargle technique on the polarization data, and its derivatives, obtained by the TBL over the last 
%10 years, it is worth putting them in the context of the periods found traditionally on light curves over the same period of time. Two aspects 
%are of  interest in this comparison: the behavior of the light curve before and after the great dimming.
%and the amplitudes of the peaks 
%at the main cited per%iods.

% In the literature, it is reported that before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variability_2006}. 
% This pulsation is often associated to the fundamental pressure mode. After the dimming, however this period appears to have disappeared 
% and only the period at 200 days,  associated to the first overtone of the fundamental mode, is visible since. 
% No explanations based upon pulsations have been ofered concerning this change of variability after or perhaps because of the dimming. But this 
% is a behaviour that

We have recovered the light curve of Betelgeuse in the visible from the AAVSO database for the last 10 years (Figure \ref{light curve Betelgeuse}).
Before the great dimming, the magnitude of Betelgeuse evolves on a timescale of a year, whereas after the dimming, the variability of Betelgeuse has shortened. 
It is clear from fig. \ref{light curve Betelgeuse} that the 
variability of Betelgeuse is evolving on a timescale shorter than one year. This qualitative change has been pointed out before,
before the great dimming the main period of Betelgeuse was of 400 days \citep{kiss_variability_2006}. 
This period is often associated to the fundamental pressure mode. After the dimming however, this period appears to have disappeared 
and only timescales shorter than 230 days are visible since \citep{dupree_great_2022}. 
No explanations have been offered concerning this change of variability after or perhaps because of the dimming, thus we can afford proposing alternative scenarios for the observed periodicities in the light curve. \

RSGs encounter mass loss event, where due to stellar winds, plasma is ejected in the interstellar medium. Those events have already been inferred in RSG like $\mu$Cep, where \cite{lopez_ariste_height_2023} found an excess of linear polarization beyond the limb of the star and associated it to rising plumes of plasma in the back hemisphere of the star. In the case of Betelgeuse, this kind of event happened only once or twice during the last 10 years, with no change in the variability of the star. However, since the great dimming, the variability of Betelgeuse has changed, meaning that this event was such different from the usual mass loss event, that it hustled the dynamics of the photosphere. \

Considering a scenario where the variability of Betelgeuse is mostly governed by convection, before the great dimming, the variability of the star was linked to the typical timescale on which convection occurs, that is 200 days for the small structures and roughly 350 days for the large structures. The great dimming impacted the behaviour of the photosphere, such that the largest structures do not evolve on timescale of the order of 350 days but rather 200 days, that is the smallest convective structures. We hypothesise that since the dimming, the photosphere is not in equilibrium and the largest convective structures are destroyed by the turbulent motion of the photosphere. Therefore, it is hard to identify periods after the great dimming, which \cite{dupree_great_2022} noticed. We expect to recover the 400 days variability within the next years, which corresponds to the time for the photosphere to go back to equilibrium.  
%But this 
%is a behaviour that fully concurs with what we find in the polarization data, and which cannot be related to any stellar pulsation but rather 
%convective dynamics. Beyond the actual values of the periods found in a Lomb-Scargle periodogram, we also see a correspondence between polarization 
%and the light curve in these changes in the observed periods before and after the great dimming. Such concurrence points to an explanation 
%in terms of convective dynamics also for the variability of the light curve.

%It may be argued that the periodograms presented in the previous sections and, in particular, those  of the photo-center displacements, 
%present low amplitude peaks at the referred main periods. One may wonder whether we are over-interpreting mere fluctuations of the periodograms, 
%while the peaks seen in the literature from analysis of the light curve are unmistakable. It is therefore worth to take advantage of the 
%availability of the AAVSO data to perform a Lomb-Scargle periodogram of the light curve in the same period covered by our polarization data and,
%furthermore, taking into consideration data around the dates for which TBL data is available. Periodograms  with such constraints are 
%shown in Fig.\ref{LSAAVSO}. Even when the full dataset for the period is taken into account (red line), the periods present small amplitudes, and are 
%fully and favourably comparable with the periodograms presented above. This is even more dramatic when the periodogram is made from data at the exclusive 
%dates at which TBL data is available (blue line). It stands out that the low amplitude peaks are rather due to both the sparse cadence and short span of 
%time of the TBL data series rather than to a physical absence. Actually, the present tests puts more emphasis on the importance of the 
%strong peaks seen at selected wavelengths in the LSD profiles, strong when compared to the peaks at similar periods arising from the light curve.

\begin{figure}[!h]
    \centering
    \includegraphics[width=0.5\textwidth]{Light_curve_Betelgeuse.png}
    \caption{Light curve of Betelgeuse from AAVSO in the V-band.}
    \label{light curve Betelgeuse}
\end{figure}

%\begin{figure}[!h]
%    \centering
%    \includegraphics[width=\linewidth]{Lomb-Scargle-TBL.png}
%    \caption{Lomb-Scargle periodogram from the AAVSO data between 2013 and 2023 (red line) and from the AAVSO data taken near TBL observations (blue line).}
%    \label{LSAAVSO}
%\end{figure}

% constant, which is not consistent with a pulsating scenario. The constant magnitude of Betelgeuse points toward another scenario than pulsation. Using surface 
% convection, we can fully explain the variability of Betelgeuse. The oscillations in the light curve are due to two mechanisms: surface convection and rising 
% plumes of plasma beyond the limb of the star. This last mechanism being random, it happens from time to time. The timescale on which convective cells evolve
% is between one and two years, which is also consistent with the 400 days period found in the literature.  After the dimming, the change of periodicity in the
% light curve is explained by surface convection, the random variation of brightness distribution could lead to the 200 days periodicity found in the literature. 
% The constant brightness of Betelgeuse for the last 6 months is explainable by the same random brightness distribution: the surface of Betelgeuse is calm, 
% leading to this constant brightness. In the future, we expect to see the light curve of Betelgeuse to change, maybe with another main period, but which 
% is in fact caused by surface convection. Thus, the main mechanism in the variability of Betelgeuse is surface convection, which is sufficient to explain the 
% variation in the magnitude of Betelgeuse.

%and not to radial or 
%other global pulsations.

% We recall that while the nature of the LSP is still in debate, the nature of the lower periods is quite commonly 
% attributed to  pressure modes 
% (e.g \cite{kiss_variability_2006}). The mechanism behind the pulsation of Betelgeuse is attributed to the $\kappa$-mechanism. In this 
% section, we attempt to 
% find an explanation to the 400 days periodicity of Betelgeuse. 
% As mentioned before, the LS periodogram of linear polarization shows a periodicity 
% around 400 days. Furthermore, the region where the LS power is high is around 40km/s. 
% This wavelength corresponds to most redshifted signal present in Betelgeuse.

%The 400-day period appears prominently in both the intensity and the total polarization profiles. And in both cases, it is mostly present 
%at wavelengths in the red wing, often beyond the 40 km/s value.
%We recall that in the present scenario for the formation of the polarization signal, this signal is supposed to remain between 
%-20 and +40km/s. Signals above 40km/s are uncommon and can be associated to
 %plasma rising behind the limb of Betelgeuse. 
%Such scenario has been observed in the RSG $\mu$ Cep and was introduced by \cite{lopez_ariste_height_2023} to explain the excess of 
%linear polarization beyond  the rest velocity of the star, which for Betelgeuse is fixed at +40km/s. 
%In $\mu$ Cep, such  linear polarization signal beyond the limit of the rest velocity of the star is very strong and particularly 
%common, two features that prompted a dedicated explanation. In Betelgeuse, such signals, when present, are weaker. However, 
%they are visible from time to time, with regularities that are captured by  the Lomb-Scargle periodogram at 400 days.
%Without further justification, we may propose that convective plumes in the hidden hemisphere of the star and 
%rising above the limb from time to time are responsible for the  400 days period in the light curve. When such plumes rise high enough 
%to be geometrically visible above the limb they show signals in intensity and polarization at redshifted wavelengths. They also
 %increase the emitting surface of the star and  the number of photons that we receive,
%leading to an increase of the brightness visible in the light curve. After a few weeks or months, the plume falls back to Betelgeuse, 
%disappearing from our point of view,  leading to a decrease of the brightness.

%The appearance of such high-rising plumes is not necessarily periodic. And the star can easily shift from active periods when such 
%plumes are a common feature to calmer periods when no plume is visible for months. $\mu$ Cep may be in one such active period presently,
%while Betelgeuse may have experienced recently one great such event in the direction of the Earth that, after dust formed, became the great 
%dimming, but since then no other high-rising plume has appeared, explaining the absence of the 400-day period since the great dimming. 
%From the point of view of convective dynamics we can offer an explanation of the 400-day period which not only explains the observation of 
%this period both in the light curve and in the polarization profiles, but also its presence at redshifted wavelengths and its present (and probably 
%temporary) disappearance since the great dimming.

% This is exactly what we see in the LSD profile of the linear polarization of Betelgeuse: 
% from time to time, there is a net signal above 40km/s, which is present for a few weeks, and then disappears. Since this signal is not always present, 
% it is found by the LS periodogram as a periodic event. However, it is not periodic at all: because the plumes rise at random moments, sometimes they 
% appear every years, sometime they do not. This would explain why we only see it before the dimming. After the dimming, this kind of event never happened 
% for the moment. Thus, the LS periodogram, at that precise frequency, finds that there is a periodic signal before the dimming, but not after, leading to
% a lower LS power at this frequency.

% To summarise what we have done so far: we were able to explain the variability of Betelgeuse using surface convection only. The convective cells play 
% the most important role in the variability of Betelgeuse. Our work does not need pulsation to explain its variability. This does not mean that there are 
% no pulsation in Betelgeuse. What we claim is that the most important factor in the variability of Betelgeuse is the timescale on which the convective 
% cells evolve. Pulsation could exist on Betelgeuse, but they are not the main contributor to the variability.

\section{Conclusion}
\label{Conclusion}

The main result  presented in this work is that the same periods identified in the light curve of Betelgeuse are also seen in the intensity and polarization
spectra, as measured with the TBL. Periods in the light curve have often been traditionally associated to pulsation due to pressure modes.
%But such pulsation would in no case affect the polarisation signals nor would explain the fact that the periods can be observed with stronger 
%amplitude at selected wavelengths but no others. % Pulsation appears to offer no coherent explanation to these observations, are hardly to the 
%observed changes in the variability since the great dimming.  The concurrence of observational facts, actual period values, evolution before 
%and after the dimming, the larger amplitude of the period peaks in polarizarion data, points towards a common origin for those variabilities. 
%And such origin cannot be pulsations.
But the interpretation of the linear polarization profiles is made in terms of convective structures in the atmosphere of Betelgeuse. We suggest 
that it is in these convective structures and their temporal evolution that we will find the true reason for the observed periodicities both 
in the light curve and in the spectropolarimetric observations. This is the main conclusion of this work: since we are able to recover the different periods in the linear polarization profiles, 
we associate the 200-days periodicity to convective dynamics while the 400-days periodicity may be related \textbf{to both stochastic pulsations and convective timescales of the largest granules.} 
This conclusion concurs with the scenario proposed by \cite{gray_mass_2008}, who attributed the variability of Betelgeuse to surface convection. 
%the observed periods are related 
%to convective dynamics and not to radial pulsations of the star.

Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the change of variability before and after the 
great dimming. We speculate that the great dimming event hustled the dynamics of the photosphere, and the largest granules were 
destroyed by the turbulent motions caused by the non equilibrium of the photosphere. Since this event, the variability of Betelgeuse has 
quickened  in adequacy with the timescales of  smaller convective cells, about 200-days. Were this be true, we shall expect in the 
coming future to see the 400-days variability re-appear as the dominant period, whenever  the photosphere returns to some equilibrium. 
Longer time  series of spectropolarimetric observations of Betelgeuse will be required to shed light on this.

%Although of secondary importance, we may suggest with more or less ease what phenomena are at the origin of the main periods. The 400-day period 
%appears to be strongly related to the appearance of plumes in the back hemisphere of the star rising high enough to be seen above the limb.
%Such plumes have been described in another RSG, $\mu$ Cep, and while less common or prominent, they are also present in Betelgeuse. The irregularity 
%of their presence justifies the 400-day period, which is captured by the Lomb-Scargle periodogram in the polarization signals attributed to 
%this phenomenon. The redshifted wavelengths at which this periodicity is found are justified by the plume moving away of the star and of the Earth while 
%raising in the back hemisphere, while the brightness increase is justified by %the extra surface of emitting plasma visible. The 200-day period is less 
%prominent but appears to be correlated to the typical time scales of the evolution of the convective structures themselves on the visible hemisphere. 
%The LSP origin is harder to pinpoint given the insufficient time length of our time series. One can speculate if it may be related to the 
%longest living photospheric structures, but this explanation does not adequately explain all the features associated to this period. Longer time 
%series of spectropolarimetric observations of Betelgeuse will be required to shed light on this.

\begin{acknowledgements}
    This work was supported by the "Programme National de Physique Stellaire" (PNPS) of CNRS/INSU co-funded by CEA and CNES.
    We acknowledge support from the French National Research Agency (ANR)
    funded project PEPPER (ANR-20-CE31-0002).
    We acknowledge the observers of the AAVSO International Database who provided more data than enough to produce this work.

    \end{acknowledgements}

    \bibliographystyle{aa}
    %\bibliographystyle{aa}

    \bibliography{art75}
\end{document}